---
title: "p8105_hw6_sz3213"
author: "Sitian Zhou"
date: "2023-11-18"
output: github_document
---

```{r, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
knitr::opts_chunk$set(
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
	fig.height = 6)

theme_set(theme_minimal())
```


## Problem 1

```{r homicide data cleaning}
homicide_df <-
  read_csv("data/homicide-data.csv") |> 
  mutate(
    city_state = str_c(city, ", ", state),
    resolved = as.numeric(disposition == "Closed by arrest")
  ) |> 
  filter(city_state != "Tulsa, AL" & victim_race != "Unknown") |> 
  filter(victim_race =="Black" | victim_race == "White") |> 
  mutate(victim_age = as.numeric(victim_age)) |> 
  drop_na()
```

```{r fit glm}
# fit glm
fit_logistic =
  homicide_df |> 
  filter(city_state == "Baltimore, MD") |> 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = _, family = binomial())

fit_logistic |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate),
         OR_CI_upper = exp(estimate + 1.96 * std.error),
         OR_CI_lower = exp(estimate - 1.96 * std.error)) |>
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

```{r OR for each city}
# loop over each city
results_df <-
  homicide_df |>
  nest(data = c(resolved, victim_sex), .by = city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolved ~ victim_sex, data = df, family = binomial())),
    results = map(models, broom::tidy)
  ) |> 
  select(-models) |> 
  unnest(results) |> 
  mutate(OR = exp(estimate),
         OR_CI_upper = exp(estimate + 1.96 * std.error),
         OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

results_df |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
  
```

```{r plot for OR and CI for each city}
#plots
results_df |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


From the plot, most cities have odds ratios that are smaller than 1, which suggests that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. However, Albuquerque, NM has a much higher OR compare to the rest of the cities, which implies that crimes with female victims have smaller odds of resolution.

## Problem 2

```{r weather data import}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```



```{r bootstrap}
bootstrap_results <-
  weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    summary = map(models, broom::glance),
    results = map(models, broom::tidy)) |> 
  select(-models) |> 
  unnest(summary, results) |> 
  select(id = .id, r.squared, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate) |> 
  mutate(
    log_b1b2 = log(tmin * prcp)
  ) |> 
  select(id, r.squared, log_b1b2)
```


```{r CI for r2}
bootstrap_results |> 
  ggplot(aes(x = r.squared)) + geom_density() + 
  labs(title = "Distribution of estimated r square")

LB_r = bootstrap_results |> pull(r.squared) |>  quantile(0.025)
UB_r =bootstrap_results |> pull(r.squared) |>  quantile(0.975)

c(LB_r, UB_r)|> 
  knitr::kable(digits = 3)
```

The distribution of $\hat{r^2}$ is slightly left-skewed, with the range from 0.86 to 0.96 and the mode around 0.92. The 2.5% and 97.5% quantiles of the distribution are `r LB_r |> round(3)` and `r UB_r |> round(3)`, respectively, and the corresponding 95% confidence interval is (`r LB_r |> round(3)`, `r UB_r |> round(3)`).


```{r CI for logb1b2}
bootstrap_results |> 
  filter(log_b1b2 != "NaN") |> 
  ggplot(aes(x = log_b1b2)) + geom_density() +
  labs(title = "Distribution of estimated log(beta1 * beta2)")

LB_b = bootstrap_results |> filter(log_b1b2 != "NaN") |> pull(log_b1b2) |>  quantile(0.025)
UB_b =bootstrap_results |> filter(log_b1b2 != "NaN") |> pull(log_b1b2) |>  quantile(0.975)

c(LB_b, UB_b)|> 
  knitr::kable(digits = 3)
```


The distribution of $log(\hat{\beta_1} * \hat{\beta_2})$ is heavily left-skewed, with the range from -12 to -4 and the mode around -5.5. The 2.5% and 97.5% quantiles of the distribution are `r LB_b |> round(3)` and `r UB_b |> round(3)`, respectively, and the corresponding 95% confidence interval is (`r LB_b |> round(3)`, `r UB_b |> round(3)`).


## Problem 3

```{r data import}
bw_df <-
  read_csv("data/birthweight.csv", na = c("", "."))
```

data cleaning

```{r data cleaning}
bw_df_clean <-
  bw_df |> 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  ) |> 
  select(bwt, everything())
```


```{r fit the model}
hypo_model = lm(bwt ~ bhead + blength + delwt + gaweeks + malform + smoken , data = bw_df_clean)
summary(hypo_model) |> broom::tidy()
# because malform has p-value > 0.05, we remove it from the model
my_model = lm(bwt ~ bhead + blength + delwt + gaweeks + smoken , data = bw_df_clean)
summary(my_model) |> broom::tidy()
```

I first hypothesized factors that are likely contributing to children's birth weights as the primary model. I included six predictors: `bhead`, `blength`, `delwt`, `gaweeks`, `malform`, and `smoken`. Then I performed the backword elimination to examine if all predictors are significant. The coefficient of `malform` variable has the greatest p-value which is 0.61, and this value is greater than the threshold of $\alpha$ equals 0.05, so I excluded it from the model and refit the rest predictors. The re-fitted model have five predictors and all coefficients have p-values less than 0.05. Thus, I will keep all predictors and conclude that the final model includes `bhead`, `blength`, `delwt`, `gaweeks`, and `smoken`. 


```{r residuals - fitted values}
bw_df_clean |> 
  add_predictions(my_model) |> 
  add_residuals(my_model) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(title = "Model residuals against fitted values")
```

```{r cross validation}
# cv
model_1 = lm(bwt ~ blength + gaweeks, data = bw_df_clean)
model_2 = lm(bwt ~ bhead + blength + babysex + 
               bhead * blength + bhead * babysex + blength * babysex, data = bw_df_clean)
cv_df <-
  crossv_mc(bw_df_clean, 100) 

cv_df <-  
  cv_df |> 
  mutate(
    my_model  = map(train, \(df) lm(bwt ~ bhead + blength + delwt + gaweeks + smoken, 
                                    data = df)),
    model_1  = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    model_2  = map(train, \(df) lm(bwt ~ bhead + blength + babysex + 
               bhead * blength + bhead * babysex + blength * babysex, data = df))) |> 
  mutate(
    rmse_my_model = map2_dbl(my_model, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_model_1 = map2_dbl(model_1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_model_2 = map2_dbl(model_2, test, \(mod, df) rmse(model = mod, data = df)))
```


```{r violin plot}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse, fill = model)) + geom_violin()
```

According to the violin plot, my_model has the lowest rmse, and model_2 has a slightly higher rmse than my_model. Model_1 has a much higher rmse compared to the other models, which implies the predictions made using this model could be less accurate than the other two.





