---
title: "p8105_hw6_sz3213"
author: "Sitian Zhou"
date: "2023-11-18"
output: github_document
---

```{r, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
knitr::opts_chunk$set(
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
	fig.height = 6)

theme_set(theme_minimal())
```


## Problem 1

```{r}
homicide_df <-
  read_csv("data/homicide-data.csv") |> 
  mutate(
    city_state = str_c(city, ", ", state),
    resolved = as.numeric(disposition == "Closed by arrest")
  ) |> 
  filter(city_state != "Tulsa, AL" & victim_race != "Unknown") |> 
  filter(victim_race =="Black" | victim_race == "White") |> 
  mutate(victim_age = as.numeric(victim_age)) |> 
  drop_na()
```

```{r}
# fit glm
fit_logistic =
  homicide_df |> 
  filter(city_state == "Baltimore, MD") |> 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = _, family = binomial())

fit_logistic |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate),
         OR_CI_upper = exp(estimate + 1.96 * std.error),
         OR_CI_lower = exp(estimate - 1.96 * std.error)) |>
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

```{r}
# loop over each city
results_df <-
  homicide_df |>
  nest(data = c(resolved, victim_sex), .by = city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolved ~ victim_sex, data = df, family = binomial())),
    results = map(models, broom::tidy)
  ) |> 
  select(-models) |> 
  unnest(results) |> 
  mutate(OR = exp(estimate),
         OR_CI_upper = exp(estimate + 1.96 * std.error),
         OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

results_df |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
  
```

```{r}
#plots
results_df |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Problem 2

```{r weather data import}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```



```{r bootstrap}
bootstrap_results <-
  weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    summary = map(models, broom::glance),
    results = map(models, broom::tidy)) |> 
  select(-models) |> 
  unnest(summary, results) |> 
  select(id = .id, r.squared, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate) |> 
  mutate(
    log_b1b2 = log(tmin * prcp)
  ) |> 
  select(id, r.squared, log_b1b2)
```


```{r CI for r2}
bootstrap_results |> 
  ggplot(aes(x = r.squared)) + geom_density() + 
  labs(title = "Distribution of estimated r square")

LB_r = bootstrap_results |> pull(r.squared) |>  quantile(0.025)
UB_r =bootstrap_results |> pull(r.squared) |>  quantile(0.975)

c(LB_r, UB_r)|> 
  knitr::kable(digits = 3)
```

The distribution of $\hat{r^2}$ is slightly left-skewed, with the range from 0.86 to 0.96 and the mode around 0.92. The 2.5% and 97.5% quantiles of the distribution are `r LB_r |> round(3)` and `r UB_r |> round(3)`, respectively, and the corresponding 95% confidence interval is (`r LB_r |> round(3)`, `r UB_r |> round(3)`).


```{r CI for logb1b2}
bootstrap_results |> 
  filter(log_b1b2 != "NaN") |> 
  ggplot(aes(x = log_b1b2)) + geom_density() +
  labs(title = "Distribution of estimated log(beta1 * beta2)")

LB_b = bootstrap_results |> filter(log_b1b2 != "NaN") |> pull(log_b1b2) |>  quantile(0.025)
UB_b =bootstrap_results |> filter(log_b1b2 != "NaN") |> pull(log_b1b2) |>  quantile(0.975)

c(LB_b, UB_b)|> 
  knitr::kable(digits = 3)
```


The distribution of $log(\hat{\beta_1} * \hat{\beta_2})$ is heavily left-skewed, with the range from -12 to -4 and the mode around -5.5. The 2.5% and 97.5% quantiles of the distribution are `r LB_b |> round(3)` and `r UB_b |> round(3)`, respectively, and the corresponding 95% confidence interval is (`r LB_b |> round(3)`, `r UB_b |> round(3)`).


## Problem 3

```{r data import}
bw_df <-
  read_csv("data/birthweight.csv", na = c("", "."))
```
